Plant Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging

### 2022-11-11 script/manuscript_MEE.Rmd
unstupiddid some paths, added more citations for spatial

---
title: Plant Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging
  Patterns of Queen Bumble Bees in the Southern Rocky Mountains, U.S.A.
author: |
  | Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E. Ogilvie $^3$, Emily J. Woodworth $^1$$^,$$^2$, Sophie Taddeo $^1$$^,$$^2$, Paul J. CaraDonna $^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant $^1$$^,$$^2$ | $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022 USA,  $^2$Plant Biology and Conservation, Northwestern University, Evanston, IL 60208, USA, $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, CO 81224, USA
abstract: |
  an abstract will be written to fill this space.
output:
  pdf_document: default
  toc: no
csl: ../citations/methods-in-ecology-and-evolution.csl
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
header-includes:
- \usepackage{endfloat}
- \usepackage{setspace}\doublespacing
- \usepackage{lineno}
- \linenumbers
---

```{r Required Libraries, message=F, warning=F, results='hide', echo = F}
library(here)
library(tidyverse)
library(bookdown)
library(knitr)
library(kableExtra)
```

```{r, eval = F, echo = F}
set_here('../')

citation('usdm')
print(citation('iNEXT'), bibtex=TRUE)
```

@@

## Study System
Observations and sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5" N, 106°59.3" W (WGS 84), 2900 m.a.s.l.), Colorado, USA (see...*APPENDIX 1* for site information). Pollinator observations of *Bombus* Latreille spp. (Apidae Latreille) were conducted from June - August of 2015. This area is characterised by high-montane/subalpine Parkland vegetation communities.

### Spatial Analyses
To develop an ecologically relevant list of vascular plant species , with expected biotic pollination, which may be present at the study sites all records adjacent to the field site were downloaded from BIEN (@bien2022), and these taxa had Species Distribution Models generated to infer their suitability. This list of species served as a reference for which species to include in the Kraken2 database.

@@

Species had all records from BIEN within a 50km border of the Omernik level 3 ecoregion which the site is located in (21. Southern Rockies), downloaded (n = 23,919), (@omernik1987ecoregions). These records were split into two sets, one for generating machine learning models, and the other for GLM and GAM models. The set for generating GLM and GAM records was thinned to reduce spatial autocorrelation in the dataset. To both datasets an additional 4029 plots collected from a random stratification of 19% of the land cover in the area of analysis were searched to create true absences (BLM CITATION- need appropriate format for journal). To achieve a larger absence dataset 1000 pseudo-absence records were generated for each taxon, each of which was greater than 10km from an occurrence record. For ML models, these pseudo-absences were reduced so that the number of presence to absence records were 1:1; to achieve this, absence records inside of 10% of the mean sample value, of the presence records, for any predictor were removed.

@@

```{r Results for distance species search, echo = F, message = F, warning = F}
citation("caret")
```

@@

logR_dat <- read.csv('/hdd/yun master/MEE_manuscript/data/PA_records_logistic_regression.csv')[,2:4] %>% 
  filter(occurrence == 0) 
# summary(logR_dat$n)
# sum(logR_dat$n)

dat <- read.csv('/hdd/yun master/MEE_manuscript/data/evaluation_table.csv')[,2:3]
data.frame(Metric  = dat[1:5,1],
                   Value   = dat[1:5,2],
                   Metric1 = dat[6:10,1],
                   Value1  = dat[6:10,2]
                   ) %>% 
  knitr::kable(caption = "Logistic regression assessing accuracy of SDMs",
    col.names = c('Metric', 'Value', 'Metric', 'Value'),
    booktabs = T) #%>% 
  #kableExtra::kable_styling(position = "float_right")
```

```{r Results for SA Logistic Regression Contingency table, echo = F, message = F, warning = F}
read.csv('/hdd/yun master/MEE_manuscript/data/Pa_conting_table.csv') %>% 
    knitr::kable(caption = "SDN evaulation contingency table",
    col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
    booktabs = T) %>%
    add_header_above(c(" " = 1, "Training" = 2, "Testing" = 2))

>>>>

---
title: Plant Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging
  Patterns of Queen Bumble Bees in the Southern Rocky Mountains, U.S.A.
author: |
  | Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E. Ogilvie $^3$, Emily J. Woodworth $^1$$^,$$^2$, Sophie Taddeo $^1$$^,$$^2$, Paul J. CaraDonna $^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant $^1$$^,$$^2$ | $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022 USA,  $^2$Plant Biology and Conservation, Northwestern University, Evanston, IL 60208, USA, $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, CO 81224, USA
abstract: |
  an abstract will be written to fill this space.
output:
  pdf_document: default
  toc: no
csl: ../citations/methods-in-ecology-and-evolution.csl
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
header-includes:
- \usepackage{endfloat}
- \usepackage{setspace}\doublespacing
- \usepackage{lineno}
- \linenumbers
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r Required Libraries, message=F, warning=F, results='hide', echo = F}
library(here)
library(tidyverse)
library(bookdown)
library(knitr)
suppressWarnings(library(kableExtra))

#print(citation('spdep'), bibtex=TRUE)
```

@@

## Study System
Observations and sample collection was conducted at The Rocky Mountain Biological Laboratory (RMBL; 38°57.5" N, 106°59.3" W (WGS 84), 2900 m.a.s.l.), Gunnison County, Colorado, USA (see...*APPENDIX 1* for site information). Pollinator observations of *Bombus* Latreille spp. (Apidae Latreille) were conducted from June - August of 2015. The six study sites are in areas characterised by high-montane/subalpine Parkland vegetation communities.

### Spatial Analyses
To develop an ecologically relevant list of vascular plant species, with expected biotic pollination, which may be present at the study sites all records adjacent to the field site were downloaded from BIEN (@bien2022), and these taxa had Species Distribution Models generated to infer their suitability. This list of species served as a reference for which species to include in the genomic sequence databases.

@@

Species had all records from BIEN within a 50km border of the Omernik level 3 ecoregion which the site is located in (No. 21' Southern Rockies'), downloaded (n = 23,919), (@omernik1987ecoregions). These records were split into two, initially identical, sets, one for generating machine learning models, and the other for Generalised Linear (GLM) and Generalized Additive Models (GAM). The set for generating GLM and GAM records was thinned to reduce spatial autocorrelation in the dataset, as measured by Morans Index (@moran1950notes, @spdep). To both datasets an additional 4029 plots collected from a random stratification of 19% of the land cover in the area of analysis were searched to create true absences (BLM CITATION- need appropriate format for journal). To achieve a larger absence dataset 1000 pseudo-absence records were generated for each taxon, each of which was greater than 10km from an occurrence record. For ML models, these pseudo-absences were reduced so that the number of presence to absence records were 1:1; to achieve this, absence records inside of 10% of the mean sample value, of the presence records, for any predictor were removed.

@@

@@

logR_dat <- read.csv('../data/PA_records_logistic_regression.csv')[,2:4] %>% 
  filter(occurrence == 0) 
# summary(logR_dat$n)
# sum(logR_dat$n)

dat <- read.csv('../data/evaluation_table.csv')[,2:3]
data.frame(Metric  = dat[1:5,1],
                   Value   = dat[1:5,2],
                   Metric1 = dat[6:10,1],
                   Value1  = dat[6:10,2]
                   ) %>% 
  knitr::kable(caption = "Logistic regression assessing accuracy of SDMs",
    col.names = c('Metric', 'Value', 'Metric', 'Value'),
    booktabs = T) #%>% 
  #kableExtra::kable_styling(position = "float_right")
```

```{r Results for SA Logistic Regression Contingency table, echo = F, message = F, warning = F}
read.csv('../data/Pa_conting_table.csv') %>% 
    knitr::kable(caption = "SDM evaluation contingency table",
    col.names = c('', ' Absence ', '  Presence  ', ' Absence ', ' Presence '), align = "lrrrr",
    booktabs = T) %>%
    add_header_above(c(" " = 1, "Training" = 2, "Testing" = 2))
